AAnalyzing Student Use of Spacing and Interleaving Strategies in Interactions with GenAI-Powered Chatbots in Programming Courses

### 2025-01-15 CSEDU2025/Paper.tex
update paper

\begin{figure*}[htbp]
  \centering
  \includegraphics[scale=0.45]{img/figure2.png}
  \caption{Examples of conversations.}
  \label{fig:graph2}
\end{figure*}

@@

For the analysis of the spacing, we initially considered a minimum interval of
60 minutes as a reference to define the significant spacing \citep{Gadella24}.
Based on this, our results showed that, of the 81 conversations analyzed, only
27 (33.3\%) had at least one message with an interval greater than 60 minutes.
Figure 2 illustrates the elapsed time between messages (delta). For example,
in conversation 15, the first message was sent at time zero, while the second
message was sent by the student after a delta of 0.6 minutes. Additionally,
message 4 in conversation 15, as shown in Figure \ref{fig:graph2}, presents a
delta of 69.1 minutes, which, according to our reference, characterizes
significant spacing. Figure \ref{fig:graph3} also shows the number of spacings
in each conversation. Among conversations that exhibited spacing, the
average number of subsections per conversation was 2.2. Therefore, a
conversation with one instance of spacing is typically divided into two distinct
study subsections.

This data suggests that there is indeed little spacing between study sessions.
It is worth noting that, in order to create a new section in CharlieBot,
students must close the browser tab and start a new conversation, which results
in losing the previous chat history.

@@

The objective of this research question was to study the time a student takes to
re-engage with the bot after sending a message for each category. This analysis
can offer insights into students' behavior: faster interactions might reflect
less reflection, whereas longer pauses could indicate the opposite.

@@

To analyze the interweaving of topics, sections were initially divided into
those that exhibited spacing and those that did not.

In conversations with the chatbot that did not include spacing (66.7\% of the total
conversations), approximately 19.2\% of the messages resulted in a change of
topic. Qualitative analysis of what led to this interweaving indicates that
part of it was due to students' requests for assistance in solving exercises
from different topics, suggesting a less active learning approach, as the time
interval for this type of interaction tends to be shorter than average.
Additionally, some instances of interweaving were observed after a prolonged
period of 20 to 50 minutes, still below the 60-minute threshold established as
the cutoff for considering spacing.

In interactions with the chatbot that included spacing (33.3\% of the total
conversations), approximately 31.4\% of the messages switched topics.
Qualitative analysis of these interactions reveals that a significant portion
of this interweaving was caused by spacing. In other words, some students use
the chatbot for quick consultations, asking a question and then returning hours
later with another question on a different subject.

Thus, the data revealed that the existing interweaving is limited and does not
suggest an effective study strategy, as it mainly arises from requests for
exercise resolution and from the natural spacing pattern of a consultation
interaction.

>>>>

\begin{figure*}[htbp]
  \centering
  \includegraphics[scale=0.42]{img/figure2.png}
  \caption{Examples of conversations.}
  \label{fig:graph2}
\end{figure*}

@@

For the spacing analysis, we initially considered a minimum interval of 60
minutes as a reference to define the significant spacing \citep{Gadella24}.
Based on this, our results showed that, of the 81 conversations analyzed, just
27, or around 34\%, had at least one message with an interval more significant
than 60 minutes. Figure \ref{fig:graph2} illustrates the elapsed time between
messages (delta). For example, in conversation 15, the first message was sent
at time zero, while the student sent the second message after a delta of 0.6
minutes. Additionally, message 4 in conversation 15, as shown in Figure
\ref{fig:graph2}, presents a delta of 69.1 minutes, which, according to our
reference, characterizes spacing. The average number of subsections per
conversation was 2.2 among conversations that exhibited spacing. Therefore, a
conversation with one spacing is divided into two distinct study sessions.

This data suggests that there is indeed little spacing between study sessions.
To create a new section in CharlieBot; students must close the browser tab and
start a new conversation, which results in losing the previous chat history.

@@

The objective of this research question was to study the time a student takes to
re-engage with the bot after sending a message for each category. This analysis
can offer insights into students' behavior: faster interactions might with the
chatbot might suggest a more passive learning posture.

@@

Switching between topics can help students identify distinct concepts inside
problems. The results show that approximately 50.8\% of the conversations
include interleaving; consequently, 49.2\% are related to a single study topic.
To gain a deeper comprehension of this interleaving, conversations were analyzed
 by separating those without spacing from those that included spacing.

Conversations without spacing comprise around 66\% of the total, with 34.2\%
showing a topic change, indicating interleaving (equivalent to 22.8\% of all
conversations). Qualitative analysis of these conversations without spacing
revealed that about 7\% of the conversations presented interleaving caused by
students' at least one request to solve quizzes. However, the speed with which
students interact with the chatbot to solve quizzes may suggest a more
superficial reflection on the answers received (see section 4.2.1). On the other
hand, 8.8\% of the conversations showed no spacing but included topic changes,
suggesting that students effectively might employ a metacognitive interleaving
strategy to distinguish concepts around a problem. Finally, 7\% of the
conversations featured interaction pauses lasting between 20 and 59 minutes but
fell short of the study's definition of spacing, which required pauses of over
60 minutes. This subset of conversations revealed a natural interleaving of
topics, reflecting a chatbot usage pattern that often mirrors how users search
Web pages.

Conversations with spacing comprise approximately 34\% of the total, with 84.2\%
indicating interleaving (equivalent to 28\% of all conversations). Pauses longer
than 60 minutes revealed similar behavior to those shorter than 60 minutes, with
students using the chatbot primarily for quick consultations, asking questions,
and returning at least one hour late with a query about a different study topic.
The argument that some students use the chatbot primarily as a tool for quick
queries is supported by data showing an average of 4.18 messages per study
session. Interestingly, only 6\% of the conversations involved students pausing
their interaction with the chatbot for more than 60 minutes before sending a
message related to the exact topic of study.

Therefore, although approximately 50.8\% of the conversations contain
interleaving, the data shows that this topic switching is limited in terms of
using a conscious metacognitive strategy.
